---
title: "ArchR Tutorial Notebook"
output:
  html_document:
    df_print: paged
---

```{r}
library(ArchR)
set.seed(1)
```

```{r}
addArchRThreads(threads = 4) 
```

```{r}
inputFiles <- getTutorialData("Hematopoiesis")
inputFiles
```


```{r}
addArchRGenome("Hg19")
```

Creating Arrow Files:

```{r}
ArrowFiles <- createArrowFiles(
  inputFiles = inputFiles,
  sampleNames = names(inputFiles),
  filterTSS = 4, #Dont set this too high because you can always increase later
  filterFrags = 1000, 
  addTileMat = TRUE,
  addGeneScoreMat = TRUE
)
```

```{r}
ArrowFiles
```
```{r}
doubScores <- addDoubletScores(
  input = ArrowFiles,
  k = 10, #Refers to how many cells near a "pseudo-doublet" to count.
  knnMethod = "UMAP", #Refers to the embedding to use for nearest neighbor search.
  LSIMethod = 1
)
```

```{r}
proj <- ArchRProject(
  ArrowFiles = ArrowFiles, 
  outputDirectory = "/HemeTutorial/",
  copyArrows = TRUE #This is recommened so that you maintain an unaltered copy for later usage.
)
```


```{r}
getAvailableMatrices(proj)
```

```{r}
proj <- filterDoublets(ArchRProj = proj)
```

## Dimensionality Reduction and Clustering

```{r}
proj <- addIterativeLSI(ArchRProj = proj, useMatrix = "TileMatrix", name = "IterativeLSI")

```
```{r}
proj
```

```{r}
proj <- addClusters(input = proj, reducedDims = "IterativeLSI")
```

## Visualizing in a 2D UMAP Embedding

```{r}
proj <- addUMAP(ArchRProj = proj, reducedDims = "IterativeLSI")
```



```{r}
p1 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "Sample", embedding = "UMAP")
```

```{r}
p2 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "Clusters", embedding = "UMAP")
```

```{r}
ggAlignPlots(p1, p2, type = "h")
```
## Assigning Clusters with Gene Scores
```{r}
proj <- addImputeWeights(proj)
```

Now we can overlay our marker gene scores on our 2D UMAP embedding.
```{r}
markerGenes  <- c(
    "CD34",  #Early Progenitor
    "GATA1", #Erythroid
    "PAX5", "MS4A1", "MME", #B-Cell Trajectory
    "CD14", "MPO", #Monocytes
    "CD3D", "CD8A"#TCells
  )

p <- plotEmbedding(
    ArchRProj = proj, 
    colorBy = "GeneScoreMatrix", 
    name = markerGenes, 
    embedding = "UMAP",
    imputeWeights = getImputeWeights(proj)
)
```

To plot a specific gene we can subset this plot list using the gene name.

```{r}
p$CD14
```

To plot all genes we can use cowplot to arrange the 9 different plots together. Each of these marker genes lights up the corresponding cell clusters

```{r}
#Rearrange for grid plotting
p2 <- lapply(p, function(x){
    x + guides(color = FALSE, fill = FALSE) + 
    theme_ArchR(baseSize = 6.5) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    theme(
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank()
    )
})
do.call(cowplot::plot_grid, c(list(ncol = 3),p2))
```
To save an editable vectorized version of this plot, we use the plotPDF() function.

```{r}
plotPDF(plotList = p, 
    name = "Plot-UMAP-Marker-Genes-W-Imputation.pdf", 
    ArchRProj = proj, 
    addDOC = FALSE, width = 5, height = 5)
```

